---
title: "NBW social clicks screening protocol"
output: html_notebook

---

```{r libraries }
library(tuneR)
```


1. I went through the field annotations in PamGuard and manually flagged those that indicated the presence of click patterns or raspberries.
Raspberries were noted either when we heard them above the water, or also when they were listened through on the hydrophone. I used the comment time stamp to identify the appropriate wave file.  

```{r}
# 1. read in field comments table -----
comm <- read.csv("/Users/anacristinaeguigurenburneo/Library/Mobile Documents/com~apple~CloudDocs/PhD/Bottlenose_Whales/NBW_Social_Clicks/raw_data/NBW_aug_2021_listening_forms.csv", header = T)

#get only flagged comments
flagged_comm<-comm[which(comm$Potential_for_screening=="yes"),]

#fix date time format
flagged_comm$UTC_posix <- as.POSIXct(flagged_comm$UTC,
                                     format = "%Y-%m-%d %H:%M:%S",
                                     tz = "UTC")


# 2. obtain recording file names----
recs<-list.files(path = "/Volumes/NBW_audio/NBW_2021_08_audio_backup_b", recursive = T, full.names = T)

recs<- as.data.frame(recs)

# 3. obtain recording file timestamps----

# extract only file names (excluding folder names)
recs$wav_names <- sub(".*/", "", recs$recs)

x <- recs$wav_names

date_time_info <- regmatches(x, gregexpr("[[:digit:]]+", recs$wav_names)) 

#create new columns for dates
recs$year   <- NA
recs$month  <- NA
recs$day    <- NA
recs$hour   <- NA
recs$minute <- NA
recs$seconds <- NA


for(i in seq_along(recs$wav_names)){
  recs$year[i]<-substr(date_time_info[[i]][2], 1,4)
  recs$month[i]<-substr(date_time_info[[i]][2], 5,6)
  recs$day[i]<-substr(date_time_info[[i]][2], 7,8)
  
  recs$hour[i]<-substr(date_time_info[[i]][3], 1,2)
  recs$minute[i]<-substr(date_time_info[[i]][3], 3,4)
  recs$seconds[i]<-substr(date_time_info[[i]][3], 5,6)  
}

# convert to a date-time object

recs$Date <- paste(recs$year, recs$month, recs$day, 
                        sep = '-')
recs$Time <- paste(recs$hour, recs$minute, recs$seconds, 
                        sep = ":")

recs$DateTime <- paste(recs$Date, recs$Time, sep = " ")

recs$DateTime <- as.POSIXct(recs$DateTime, 
                                 format = "%Y-%m-%d %H:%M:%S", 
                                 tz = "UTC")

# find file duration

recs$duration_min <- NA


for( i in seq_along(recs$duration_min)){
  try({readfile<-recs$recs[i]
  audio<-readWave(readfile, header=TRUE)
  recs$duration_min[i] <- round((audio$samples / audio$sample.rate)/60, 2)})
}

#find end time
recs$DateTime_end<- recs$DateTime + recs$duration_min*60



# 3. match file names to comments -----
# get files within which comments fall

# get interval information
library(lubridate)
rec_intervals        <- as.list(interval(recs$DateTime, recs$DateTime_end))


head(recs$DateTime)
i<-2

flagged_comm$wav_file_name <- NA
flagged_comm$wav_file_path <- NA

for(i in seq_along(flagged_comm$wav_file_name)){
  flagged_comm$wav_file_name[i] <- recs$wav_names[which(recs$DateTime<=flagged_comm$UTC_posix[i]&recs$DateTime_end>=flagged_comm$UTC_posix[i])] 
  flagged_comm$wav_file_path[i] <- recs$recs[which(recs$DateTime<=flagged_comm$UTC_posix[i]&recs$DateTime_end>=flagged_comm$UTC_posix[i])] 
  }




# 4. copy flagged files to computer (temporarily) ------



# 4. create a folder that contains only these wav files ------

#  place new folder path and name here

folder <- "/Users/anacristinaeguigurenburneo/Desktop/NBW_raspberry_files"



dir.create(folder)#creates a folder where all files will be stored


file.copy(flagged_comm$wav_file_path, folder)

# manually transfer them to Hard Drive, since I don't have enough USB 
# adapters


# 5. save table with observations-----

write.csv(flagged_comm, "Processed_Data/Raspberries_and_patterns_screening.csv")
```



2. Next, I excluded the files that met the following criteria from analysis:
    - Odontocetes where visually detected within 30 from each wave file (using Identify_clean_raspberries_patterns.R) 
    - Visibility was less than 1 nm.  
    
    
3. This resulted in 10 files for initial screening, which I then copied to the hard drive.   


4. Created a filtered version so I could explore on PamLab simultaneously on Matlab (Filter 1,000 - 90,0000 hz)  
